Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-05 paper/tv.tex
m

    \acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

\definecolor{darkpastelred}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{\textcolor{black}{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\newcommand\pure[1]{ \textcolor{black}{#1}}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize #1}}}}}
\newcommand{\anynotevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize $\overline{\text{#1}}$}}}}}
\newcommand{\seq}{\cdot}
\newcommand{\choice}{\vee}
\newcommand{\fullname}{Integrated Dependent Effects}
\newcommand{\name}{integrated effects}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\esn}[2]{\ensuremath{#1^{#2}}}
\newcommand{\empt}{\textcolor{black}{\ensuremath{\epsilon}}}
\newcommand{\bott}{\textcolor{black}{\ensuremath{\bot}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit}  
\newcommand{\toolName}{Tvide}
\def\defeq{\ensuremath{\,\triangleq}}
\definecolor{regalia}{rgb}{0.32, 0.18, 0.5}
\newcommand{\inclusion}{\code{\mathcal{I}}}
\definecolor{mypink}{RGB}{219, 48, 122}
\definecolor{darklavender}{rgb}{0.45, 0.31, 0.59}
\definecolor{deepcerise}{rgb}{0.85, 0.2, 0.53}
\newcommand\figref[1]{Fig. \textcolor{black}{\ref{#1}}.}
\newcommand\tabref[1]{Table \textcolor{black}{\ref{#1}}.}
\newcommand\secref[1]{Sec. \textcolor{black}{\ref{#1}}}
\newcommand{\timedL}{\code{C^{t}}}

@@

The specifications of  \codem{addSugar(n)} indicates that the 
input  \code{n} is non-negative; the method 
generates a finite trace where \anyevent{Sugar} takes a no less 
than \code{d} milliseconds delay to finish. 
Although these examples are simple, they show the benefits of deploying 
value-dependent 
time bounds, which is beyond the capability of Timed Automata. Intuitively, if traditional Timed Automata define an 
\emph{exact} transition system, \timedEffects\ define 
a set (possibly infinite) of exact transition systems.

@@

%\subsection{HipHop.js and Dependent Effects}
\subsection{\timedEffects}
As shown in \figref{fig:overview_eg1}, we define Hoare-triple style specifications (enclosed in \textcolor{darklavender}{\ttfamily{{/*@ ... @*/}}}) for each program, which leads to a compositional verification strategy, where static checking and timed temporal verification can be done locally. %The set of signals to be present in one logical time instance are represented within one \code{\_}.

@@

After emitting the signal \anyevent{Connecting}, it emits  \anyevent{Connected} if the connection succeeds before the  preemptive deadline; otherwise, no signal is emitted. 
The precondition \code{d{>}3 : \{
\}^\star \cdot\{\m{Login}\} } requires that the input variable \code{d} is greater than 3, and before entering into this module, the signal \anyevent{Login} should be emitted at the last time instance, indicating that a login request has been sent. 
The postcondition contains two parts and connected using the disjunction mark \code{{\vee}}: when \code{3{<}t{<}d}, the effects contains two time instances, \code{\{\m{Connecting}\}} and \code{\{\m{Connected}\}}, while the first  time instance finishes at time \code{t}, %(assuming to reset the clock when executing the module), 
which is no later than \code{\m{d}} milliseconds; otherwise, the effects contains one time instance \code{\{Connecting\}} which finishes at time \code{d}. \emph{(See a demo from \cite{CODE1})}

As for another example, shown in \figref{fig:overview_eg2}, the module {\ttfamily{main}} %contains a \code{\m{fork/par}   } statement, which 
spawns two threads running in  parallel.
The first thread firstly waits for the signal \anyevent{Ready}, then emits  \anyevent{Go}. 
The second thread firstly emits  \anyevent{Prep}, then calls the function {\ttfamily{cook}} asynchronously. 
The precondition of {\ttfamily{main}} requires no arithmetic constraints on its input values (expressed as \code{\m{true}}), neither any pre-traces  (expressed as \code{\m{emp}} or \code{\epsilon}, indicating an empty trace). The postcondition ensures a trace, where the second time instance contains at least one signal \code{\anyevent{Cook}} and finishes within 3 milliseconds after the completion of the first time instance. Then we do not care about the rest of the trace (\code{\_} is similar to a wildcard). 
%We omit the implementation of \code{cook} here, while provide the specifications of it. 
Taking {\ttfamily{main}} as an example, the work flow of the forward verifier is presented in the following sub-section. \emph{(See a demo from \cite{CODE2})}

@@

The effects states (1) and (4) are initial effects entering into the \code{fork/par} statement. The effects state (2) is obtained by [FV-Await], which concatenates a blocking signal 
%(with a question mark) 
%followed by a new empty time instance
 to the current effects. 
%The effects state (3) is obtained by [FV-Emit], which simply adds  %concatenates an empty time instance to the current effects. %to be the new current state. 
The effects states (3) and (5) are obtained by [FV-Emit], which simply adds the emitted  signal to the current time instance. The intermediate effects state of (7) is obtained by [FV-Call]. 
Before each function call, it invokes the TRS to check whether the current effects state satisfies the precondition of the callee module. If it is not satisfied, the verification fails, otherwise it concatenates the callee's postcondition to the current effects state. The final effects state of (7) is obtained by [FV-Async], which adds a new time instance \code{\{Ready\}} to the current effects state, indicating that the asynchronous program has been resolved. 
In step (8), we parallel compose the effects from both of the branches, and normalize the final effects. 
After these states transformations, step (9) checks the satisfiability of the inferred effects against the declared postcondition by invoking the TRS.

@@

Similarly, we defined the \defref{Inclusion} for unfolding the inclusions  between \timedEffects, where \code{D_{I\# t}(\effect)} is the partial derivative of \code{\effect} w.r.t the time instance \code{I} and time bound \code{t}.

@@

Next, we continue with the step (9) in \figref{fig:forward_example}, to 
demonstrate how the TRS handles arithmetic constraints and dependent values. 
%We use the postcondition proving for module \code{main} as an example to briefly demonstrate the rewriting process. 
As shown in \tabref{tab:rewriting_tree_send}, it automatically proves that the inferred effects of {\ttfamily{main}}
 satisfy the declared postcondition. 
%property \code{\effect^\prime  {=} 0{\leq} t {<}3 : \_\cdot(\{Cook\}\# t)\cdot\_^\star}, 
%where \code{\effect^\prime} simply indicates that the signal \anyevent{Cook} needs to happen at the second time instance, and takes no more than \code{3} milliseconds after the first time instance has completed; and \code{\_^\star} means we do not care the rest of the traces. 
We mark the rewriting rules (cf. \secref{sec:Entailment_Prover}) in \textcolor{mGray}{[gray]}.

{
\begin{table*}[ht]
\centering
      \vspace{0mm}
\caption{\label{tab:rewriting_tree_send} The inclusion proving example. (For simplicity, we use 3 seconds to replace the 3 milliseconds.)}

@@

Note that time instance \code{\{Prep\}} entails \code{\_} because the former contains more constraints. We formally define the subsumption for time instances in \defref{Subsumption}.
Intuitively, we use \code{[DISPROVE]} wherever the left-hand side (LHS) is \emph{nullable}\footnote{If the event sequence is possibly empty, i.e. contains \code{\empt}, we call it nullable, formally defined in \defref{Nullable}.} while the right-hand side (RHS) is not. 
\code{[DISPROVE]} is the heuristic refutation step to disprove the inclusion early, improving the verification efficiency.

As shown in \tabref{tab:rewriting_tree_send}, in step \code{\textcircled{1}}, we rename the time variables to avoid the name clashes between the antecedent and the consequent.  
In step \code{\textcircled{2}}, we design a new rule \code{[SPLIT]} to accommodate the extended real-time constraints, which introduces two new time variables \code{t_L^1} and \code{t_L^2}, and 
extends the previous constraint \code{t_L {<} 3} with constraints  \code{t_L^1 {+} t_L^2 {=} t_L}. Then \code{t_L^1} and \code{t_L^2} mark the real-time constraints for time instances  \code{\{Prep\}} and \code{\{Cook\}} respectively. 
Then in step \code{\textcircled{3}}, we eliminate \code{\{Prep\}} from the LHS and \code{\_} from the RHS, as the instances entailment  \code{\{Prep\} \subseteq \_} holds.  
In step \code{\textcircled{4}}, in order to conduct a further unfolding, we unify time variables \code{t_L^2} and \code{t_R} by adding the constraint \code{t_L^2 {=} t_R}. 
In step \code{\textcircled{5}}, from the RHS, since \code{\_^\star {=} emp \vee \_ \cdot \_^\star}, eliminating one time instance \code{\{Ready\}} from it will lead to \code{\bot \vee \_^\star}, which can be normalised into \code{\_^\star} in step \code{\textcircled{6}}.
At the end of the rewriting, we manage to prove that \code{(t_L {<}3 {\wedge} t_L^1 {+} t_L^2 {=}t_L {\wedge}t_L^2 {=} t_R)  \Rightarrow  t_R {<} 3} \footnote{The proof obligations generated by the verifier are discharged using constraint solver Z3 \cite{de2008z3}.}; therefore, the proof succeed.

@@

A transition of the system is of the form $c \xrightarrow[]{l} c^\prime$ where $c$ and $c^\prime$ are 
the system configurations before and after the transition respectively. 
Transition labels \code{l} consist: 
%We adopt the following naming convention for transition labels: 
$t$, denoting a non-negative real integer; 
$\tau$, denoting an invisible event;
$\anyevent{A} \in \Sigma $ is an observable 
event; and $\sigma \in \Sigma\cup \{\tau\}$. 
For instance, $c \xrightarrow[]{\text{t}} c^\prime$ denotes a transition of \codem{t} time units elapsing. 
In the following, we present the firing rules which are associated with 
the timed process constructs.

@@

    \m{(Timed\ Effects)} &  \effect &\code{\  ::=\ }&
    \code{
          (\pure{\pi}, \es)
   \ | \ \effect_1 \vee \effect_2
}
    \\
 \m{(Timed \ Sequence)}  & \ \code{\es}   &\code{\ ::=\ }&

@@

\\
\m{(Events)} & ev  & \ ::= \ & 
\code{
  \anyevent{A}(v, \pi, \alpha^*)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

@@

    \m{(Pure)}  & \code{\code{\pure{\pi}}}&\code{\  ::=\ }&
\pure{ \pure{True}}
   \ | \  \pure{False}
   \ | \ \pure{A(}{t_1, t_2}\pure{)}
  \   | \  \pure{\pure{\pi_1}} \wedge \code{\pure{\pi}}_2
   \  | \ \pure{\pure{\pi_1}} \vee \code{\pure{\pi}}_2
   \  | \ \neg\code{\pure{\pi}}

@@

    & &&\code{
       | \  \pure{\pure{\pi_1}} \Rightarrow \pure{\pure{\pi_2}}
   \  | \ \forall x. \code{\pure{\pi}}
   \  | \ \exists x. \code{\pure{\pi}}
   }\\

    \m{(Real\text{-}Time\ Term)}  & \code{t} &\code{\  ::=\ }&
    \code{ c
    \ | \  x
    \ | \  t_1{+}t_2
    \ | \  t_1\text{-}t_2
    %\ | \ k {\times} n
   }

    \\~\\
    \multicolumn{4}{c}{
     % \code{(Events)  \anyevent{A} ::\in \Sigma}   
   %\qquad
    \code{c ::\in \mathbb{Z}^{+}} 
     \qquad 
           \code{x  ::\in \textbf{var} } 
           \qquad

     \m{(Real\ Time\ Bound)}\ \code{\#}
     \qquad
        \m{(Kleene\ Star) \ \star}

@@

The  syntax of the \timedEffects is formally defined in \figref{fig:Syntax_of_Types_and_Effects} Effects is  a conditioned time  instance sequence \code{\pure{\pi} : \es} or a disjunction of two effects $\effect_1 \vee \effect_2$.
Timed sequences comprise \textit{nil} ($\bott$);
 an empty trace $\empt$;
a single time instance represented by \code{I};
a waiting for a single signal \code{\textbf{S}?};
  %where \anyevent{a} is an element from a finite set of events, donated by \code{\Sigma};
%parametrized event \code{ \anyevent{a}.\iota} where \code{\iota} is a list pattern expressing a possibly infinite list (for example, \code{\anyevent{a}.[1,2,3,4]},  \code{\anyevent{a}.[1 .. 4]} and \code{\anyevent{a}.[1, 2.. 4]} all refer to the finite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot \anyevent{a}.4},
%while \code{\anyevent{a}.[1 ..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot}..., and \code{\anyevent{a}.[1, 3..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.3 \cdot \anyevent{a}.5\cdot}...);
 sequences concatenation \code{\es_1\cdot \es_2};
disjunction  \code{\es_1\vee \es_2};
synchronous parallelism   \code{\es_1 || \es_2}.
%negation \code{\neg \es};
 %

 We introduce a new operator \code{\#}, and the effects \code{\es \# t } represents that a trace takes real-time \code{t} to complete, where \code{t} is a \emph{term}. 
A time sequence can be constructed by \code{\star}, representing zero or more times repetition of a trace.

@@

\begin{figure}[ht]
    \vspace{-1mm}
    \renewcommand{\arraystretch}{1}
\begin{align*} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          Disjunction                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \effect_1 {\vee} \effect_2}  
&\m{iff}\quad 
&  d, \varphi \models \effect_1 \ or \ d, \varphi \models \effect_2
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          EMPTY                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models \pi : \empt }  
&\m{iff}\quad 
&  d{=}0 \ and\  {\tt{SAT}(\pi)}  \ and \ \varphi {=} [] 
\\[0.2em]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                          I                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : ev }  &\m{iff}\quad 
& d {\geq} 0 \ and\ {\tt{SAT}(\pi)}  \ and \ 
\varphi {=} \code{[ev]}
  \\[0.2em]
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            CONCATE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1{\cdot}\es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \exists \code{\varphi_1, \varphi_2, d_1, d_2}\ and\ \code{
\varphi {=} \varphi_1 {+}{+} \varphi_2, d {=} d_1 {+} d_2 \ and 
}\\
& & &
\code{d_1, \varphi_1 \models  \pi : \es_1}  \ and \ 
\code{d_2, \varphi_2 \models  \pi : \es_2}  
 \\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       DISJUNCTION                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : (\es_1 {\vee} \es_2)}  
&\m{iff}\quad 
& {\tt{SAT}(\pi)}  \ and \ \code{d, \varphi \models  \pi :  \es_1 } \ or \ 
 \code{d, \varphi \models  \pi :  \es_2}  
\\[0.2em]
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                       timed                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d, \varphi \models  \pi : \es \# t }  
&\m{iff}\quad 
& d, \varphi \models  (\pi \wedge t{=}d) : \es
\\[0.2em]
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            KLEENE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
&\code{d,  \varphi \models \pi : \esn{\es}{\star}  }  
&\m{iff}\quad 
&d, \varphi \models \pi :  \empt \ or \ 
d, \varphi \models \pi :  (\es \cdot \esn{\es}{\star} ) 
 \\[0.2em]
   %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                            FALSE                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
& \code{d, \varphi \models False : \bot }  
&\m{iff}\quad 
&   otherwise
%{\tt{UNSAT}(\pi)}  \ or \  \varphi  {=} \bot
%\exists t, d(t) \not\models \llbracket \pure{\pi} \rrbracket_{t}
\end{align*}
    \vspace{-3mm}
\caption{Semantics of \timedEffects.}
\label{fig:Sementic}
  \vspace{-1mm}
\end{figure}

>>>>

    \acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}

\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

\definecolor{darkpastelred}{rgb}{0, 0, 0}
\definecolor{backgroundColour}{rgb}{0.95,0.95,0.92}
\definecolor{brickred}{rgb}{0.8, 0.25, 0.33}
\definecolor{red}{rgb}{0.6,0,0} 
\definecolor{blue}{rgb}{0,0,0.6}
\definecolor{green}{rgb}{0,0.8,0}
\definecolor{cyan}{rgb}{0.0,0.6,0.6}
\definecolor{cloudwhite}{rgb}{0.9412, 0.9608, 0.8471} 
\definecolor{davysgrey}{rgb}{0.33, 0.33, 0.33}
\definecolor{deepfuchsia}{rgb}{0.76, 0.33, 0.76}
\definecolor{deeplilac}{rgb}{0.6, 0.33, 0.73}
\definecolor{deepskyblue}{rgb}{0.0, 0.75, 1.0}
\definecolor{indianred}{rgb}{0.8, 0.36, 0.36}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{darkred}{rgb}{0.55, 0.0, 0.0}
\usepackage{wrapfig}
\newcommand{\effect}{\textcolor{black}{\ensuremath{\mathrm{\Phi}}}}
\definecolor{color_pure}{RGB}{3,35,14}
\newcommand\pure[1]{ \textcolor{black}{#1}}
\usepackage{epigraph}
\newcommand{\anyevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize #1}}}}}
\newcommand{\anynotevent}[1]{{\textcolor{darkred}
{{\textbf{\footnotesize $\overline{\text{#1}}$}}}}}
\newcommand{\seq}{\cdot}
\newcommand{\choice}{\vee}
\newcommand{\fullname}{Integrated Dependent Effects}
\newcommand{\name}{integrated effects}
\newcommand{\code}[1]{{\tt{\ensuremath{\m{#1}}}}}
\newcommand{\codeme}[1]{{\tt{\ensuremath{#1}}}}
\newcommand{\esn}[2]{\ensuremath{#1^{#2}}}
\newcommand{\empt}{\textcolor{black}{\ensuremath{\epsilon}}}
\newcommand{\bott}{\textcolor{black}{\ensuremath{\bot}}}
\newcommand{\CONTAIN}{\sqsubseteq}
\newcommand{\underscore}{\textcolor{vividauburn}{\code{\_}}}
\newcommand{\m}{\mathit} 
\newcommand{\lappend}{\mathrel{\texttt{++}}}

\newcommand{\toolName}{Tvide}
\def\defeq{\ensuremath{\,\triangleq}}
\definecolor{regalia}{rgb}{0.32, 0.18, 0.5}
\newcommand{\inclusion}{\code{\mathcal{I}}}
\definecolor{mypink}{RGB}{219, 48, 122}
\definecolor{darklavender}{rgb}{0.45, 0.31, 0.59}
\definecolor{deepcerise}{rgb}{0.85, 0.2, 0.53}
\newcommand\figref[1]{Fig. \textcolor{black}{\ref{#1}}.}
\newcommand\tabref[1]{Table \textcolor{black}{\ref{#1}}.}
\newcommand\secref[1]{Sec. \textcolor{black}{\ref{#1}}}
\newcommand{\timedL}{\code{C^{t}}}

@@

The specifications of  \codem{addSugar(n)} indicates that the 
input  \code{n} is non-negative; the method 
generates a finite trace where \anyevent{Sugar} takes a no less 
than \code{d} time units delay to finish. 
Although these examples are simple, they show the benefits of deploying 
value-dependent 
time bounds, which is beyond the capability of Timed Automata. Intuitively, if traditional Timed Automata define an 
\emph{exact} transition system, \timedEffects\ define 
a set (possibly infinite) of exact transition systems.

@@

%\subsection{HipHop.js and Dependent Effects}
\subsection{\timedEffects}
As shown in \figref{fig:overview_eg1}, we define Hoare-triple style specifications (enclosed in \textcolor{darklavender}{\ttfamily{{/*@ ... @*/}}}) for each program, which leads to a compositional verification strategy, where static checking and timed temporal verification can be done locally. %The set of signals to be present in one logical event are represented within one \code{\_}.

@@

After emitting the signal \anyevent{Connecting}, it emits  \anyevent{Connected} if the connection succeeds before the  preemptive deadline; otherwise, no signal is emitted. 
The precondition \code{d{>}3 : \{
\}^\star \cdot\{\m{Login}\} } requires that the input variable \code{d} is greater than 3, and before entering into this module, the signal \anyevent{Login} should be emitted at the last event, indicating that a login request has been sent. 
The postcondition contains two parts and connected using the disjunction mark \code{{\vee}}: when \code{3{<}t{<}d}, the effects contains two time instances, \code{\{\m{Connecting}\}} and \code{\{\m{Connected}\}}, while the first  event finishes at time \code{t}, %(assuming to reset the clock when executing the module), 
which is no later than \code{\m{d}} time units; otherwise, the effects contains one event \code{\{Connecting\}} which finishes at time \code{d}. \emph{(See a demo from \cite{CODE1})}

As for another example, shown in \figref{fig:overview_eg2}, the module {\ttfamily{main}} %contains a \code{\m{fork/par}   } statement, which 
spawns two threads running in  parallel.
The first thread firstly waits for the signal \anyevent{Ready}, then emits  \anyevent{Go}. 
The second thread firstly emits  \anyevent{Prep}, then calls the function {\ttfamily{cook}} asynchronously. 
The precondition of {\ttfamily{main}} requires no arithmetic constraints on its input values (expressed as \code{\m{true}}), neither any pre-traces  (expressed as \code{\m{emp}} or \code{\epsilon}, indicating an empty trace). The postcondition ensures a trace, where the second event contains at least one signal \code{\anyevent{Cook}} and finishes within 3 time units after the completion of the first event. Then we do not care about the rest of the trace (\code{\_} is similar to a wildcard). 
%We omit the implementation of \code{cook} here, while provide the specifications of it. 
Taking {\ttfamily{main}} as an example, the work flow of the forward verifier is presented in the following sub-section. \emph{(See a demo from \cite{CODE2})}

@@

The effects states (1) and (4) are initial effects entering into the \code{fork/par} statement. The effects state (2) is obtained by [FV-Await], which concatenates a blocking signal 
%(with a question mark) 
%followed by a new empty event
 to the current effects. 
%The effects state (3) is obtained by [FV-Emit], which simply adds  %concatenates an empty event to the current effects. %to be the new current state. 
The effects states (3) and (5) are obtained by [FV-Emit], which simply adds the emitted  signal to the current event. The intermediate effects state of (7) is obtained by [FV-Call]. 
Before each function call, it invokes the TRS to check whether the current effects state satisfies the precondition of the callee module. If it is not satisfied, the verification fails, otherwise it concatenates the callee's postcondition to the current effects state. The final effects state of (7) is obtained by [FV-Async], which adds a new event \code{\{Ready\}} to the current effects state, indicating that the asynchronous program has been resolved. 
In step (8), we parallel compose the effects from both of the branches, and normalize the final effects. 
After these states transformations, step (9) checks the satisfiability of the inferred effects against the declared postcondition by invoking the TRS.

@@

Similarly, we defined the \defref{Inclusion} for unfolding the inclusions  between \timedEffects, where \code{D_{I\# t}(\effect)} is the partial derivative of \code{\effect} w.r.t the event \code{I} and time bound \code{t}.

@@

Next, we continue with the step (9) in \figref{fig:forward_example}, to 
demonstrate how the TRS handles arithmetic constraints and dependent values. 
%We use the postcondition proving for module \code{main} as an example to briefly demonstrate the rewriting process. 
As shown in \tabref{tab:rewriting_tree_send}, it automatically proves that the inferred effects of {\ttfamily{main}}
 satisfy the declared postcondition. 
%property \code{\effect^\prime  {=} 0{\leq} t {<}3 : \_\cdot(\{Cook\}\# t)\cdot\_^\star}, 
%where \code{\effect^\prime} simply indicates that the signal \anyevent{Cook} needs to happen at the second event, and takes no more than \code{3} time units after the first event has completed; and \code{\_^\star} means we do not care the rest of the traces. 
We mark the rewriting rules (cf. \secref{sec:Entailment_Prover}) in \textcolor{mGray}{[gray]}.

{
\begin{table*}[ht]
\centering
      \vspace{0mm}
\caption{\label{tab:rewriting_tree_send} The inclusion proving example. }

@@

Note that event \code{\{Prep\}} entails \code{\_} because the former contains more constraints. We formally define the subsumption for time instances in \defref{Subsumption}.
Intuitively, we use \code{[DISPROVE]} wherever the left-hand side (LHS) is \emph{nullable}\footnote{If the event sequence is possibly empty, i.e. contains \code{\empt}, we call it nullable, formally defined in \defref{Nullable}.} while the right-hand side (RHS) is not. 
\code{[DISPROVE]} is the heuristic refutation step to disprove the inclusion early, improving the verification efficiency.

As shown in \tabref{tab:rewriting_tree_send}, in step \code{\textcircled{1}}, we rename the time variables to avoid the name clashes between the antecedent and the consequent.  
In step \code{\textcircled{2}}, we design a new rule \code{[SPLIT]} to accommodate the extended real-time constraints, which introduces two new time variables \code{t_L^1} and \code{t_L^2}, and 
extends the previous constraint \code{t_L {<} 3} with constraints  \code{t_L^1 {+} t_L^2 {=} t_L}. Then \code{t_L^1} and \code{t_L^2} mark the real-time constraints for time instances  \code{\{Prep\}} and \code{\{Cook\}} respectively. 
Then in step \code{\textcircled{3}}, we eliminate \code{\{Prep\}} from the LHS and \code{\_} from the RHS, as the instances entailment  \code{\{Prep\} \subseteq \_} holds.  
In step \code{\textcircled{4}}, in order to conduct a further unfolding, we unify time variables \code{t_L^2} and \code{t_R} by adding the constraint \code{t_L^2 {=} t_R}. 
In step \code{\textcircled{5}}, from the RHS, since \code{\_^\star {=} emp \vee \_ \cdot \_^\star}, eliminating one event \code{\{Ready\}} from it will lead to \code{\bot \vee \_^\star}, which can be normalised into \code{\_^\star} in step \code{\textcircled{6}}.
At the end of the rewriting, we manage to prove that \code{(t_L {<}3 {\wedge} t_L^1 {+} t_L^2 {=}t_L {\wedge}t_L^2 {=} t_R)  \Rightarrow  t_R {<} 3} \footnote{The proof obligations generated by the verifier are discharged using constraint solver Z3 \cite{de2008z3}.}; therefore, the proof succeed.

@@

A transition of the system is of the form $c \xrightarrow[]{l} c^\prime$ where $c$ and $c^\prime$ are 
the system configurations before and after the transition respectively. 
Transition labels \code{l} consist: 
%We adopt the following naming convention for transition labels: 
$t$, denoting a non-negative real integer; 
$\tau$, denoting an invisible event;
$\anyevent{A} \in \Sigma $ is an observable 
event; and $\sigma \in \Sigma\cup \{\tau\}$. 
For event, $c \xrightarrow[]{\text{t}} c^\prime$ denotes a transition of \codem{t} time units elapsing. 
In the following, we present the firing rules which are associated with 
the timed process constructs.

@@

    \m{(Timed\ Effects)} &  \effect &\code{  ::= }&
    \code{
          (\pure{\pi}, \es)
   \ | \ \effect_1 \vee \effect_2
}
    \\
 \m{(Timed \ Sequence)}  & \ \code{\es}   &\code{ ::=}&

@@

\\
\m{(Events)} & ev  &  ::=  & 
\code{
  \anyevent{A}(v, \pi, \alpha^*)
    \ | \ \anynotevent{A} 
    \ | \ \_  
}

@@

    \m{(Pure)}  & \code{\code{\pure{\pi}}}&\code{  ::= }&
\pure{ \pure{True}}
   \ | \  \pure{False}
   \ | \ \pure{A(}{t_1, t_2}\pure{)}
  \  |   \pure{\pure{\pi_1}} {\wedge} \code{\pure{\pi}}_2
   \ |  \pure{\pure{\pi_1}} {\vee} \code{\pure{\pi}}_2
   \ |  \neg\code{\pure{\pi}}
  \  |   \pure{\pure{\pi_1}} {\Rightarrow} \pure{\pure{\pi_2}}
   \ |  \forall x. \code{\pure{\pi}}
   \ |  \exists x. \code{\pure{\pi}}

@@

    \m{(Real\text{-}Time\ Term)}  & \code{t} &\code{  ::= }&
    \code{ c
    \ | \  x
    \ | \  t_1{+}t_2
    \ | \  t_1\text{-}t_2
    %\ | \ k {\times} n
   }

    \\~\\
    \multicolumn{4}{c}{
     % \code{(Events)  \anyevent{A} ::\in \Sigma}   
   %\qquad
    \code{c ::\in \mathbb{Z}} 
     \qquad \qquad 
           \code{x  ::\in \textbf{var} } 
           \qquad\qquad

     \m{(Real\ Time\ Bound)}\ \code{\#}
     \qquad \qquad 
        \m{(Kleene\ Star) \ \star}

@@

The  syntax of the \timedEffects is formally defined in 
\figref{fig:Syntax_of_Types_and_Effects} Effects is 
 a conditioned time  event sequence \code{\pure{\pi},  \es} or a disjunction of two effects $\effect_1 \vee \effect_2$.
Timed sequences comprise \textit{nil} ($\bott$);
 an empty trace $\empt$;
a single event \code{ev};
  %where \anyevent{a} is an element from a finite set of events, donated by \code{\Sigma};
%parametrized event \code{ \anyevent{a}.\iota} where \code{\iota} is a list pattern expressing a possibly infinite list (for example, \code{\anyevent{a}.[1,2,3,4]},  \code{\anyevent{a}.[1 .. 4]} and \code{\anyevent{a}.[1, 2.. 4]} all refer to the finite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot \anyevent{a}.4},
%while \code{\anyevent{a}.[1 ..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.2 \cdot \anyevent{a}.3 \cdot}..., and \code{\anyevent{a}.[1, 3..]} refers to the infinite event sequence \code{\anyevent{a}.1 \cdot \anyevent{a}.3 \cdot \anyevent{a}.5\cdot}...);
 sequences concatenation \code{\es_1\cdot \es_2};
disjunction  \code{\es_1\vee \es_2};
parallel composition of two processes is written as \code{\es_1 || \es_2}, where \code{es_1} and \code{es_2} 
may communicate via multi-party event synchronization or shared variables. 
We introduce a new operator \code{\#}, and the effects \code{\es \# t } represents that a trace takes real-time \code{t} to complete, where \code{t} is a \emph{term}. 
A time sequence can be constructed by \code{\star}, representing zero or more times repetition of a trace.

\code{\anyevent{A}(v, \pi, \alpha^*)} stands for an event labeled 
as \anyevent{A} with parameter \code{v}, guard \code{\pi} and assignment operations \code{\alpha^*}. 
To make the specification more expressive, events can also be 
\code{\anynotevent{A}}, representing  events which are not 
labeled using \anyevent{A}; and a wildcard \code{\_}, which matches to all the events.

@@

To define the model, 
\code{var} is the set of program variables, 
\code{val} is the set of primitive values, 
\code{\es} is the set of event sequences (or event multi-trees, per se), indicating the sequencing constraints on temporal behaviour.
Let \code{d, s, \varphi \models \effect} denote the model relation, i.e., 
the effects \code{\effect} take exactly \code{d} time units to complete; 
the stack \code{s} and linear temporal events \code{\varphi} satisfy the 
temporal effects \code{\effect}, with \code{d}, \code{s}, \code{\varphi} 
from the following concrete domains: \code{d}  {\defeq}\  \code{\mathbb{N}}, 
\code{s}  \defeq\  \code{var {\rightarrow} val } and \code{\varphi}   \defeq\ \code{[ev]} \ .

As shown in \figref{fig:Sementic}, we define the semantics of our effects. 
We use $\lappend$ to represent the append operation of two event sequences. 
We use $[]$ to represent the empty sequence, \code{[ev]} to represent 
the sequence only contains a singleton event.
